### 2016-11-10 paper.tex
Proof-read sections I and II

\documentclass{article}

%%%% Packages %%%%

\usepackage{minted} % Used for JSON highlighting
\usepackage{pdfcomment} % Used for notes and todos
\usepackage[margin=1in]{geometry} % Increases margins
\usepackage[table]{xcolor} % Color definitions
\usepackage{algorithm} % Algorithm float
\usepackage{algpseudocode} % Algorithmic environment
\usepackage{xspace}

%%%% Commands %%%%

\newcommand{\note}[2]{\pdfmargincomment[color=yellow,author=#1,open=true]{#2}}
\newcommand{\todo}[1]{\color{red}TODO: #1\color{black}}
\newcommand{\boutiques}{Boutiques\xspace}
\newcommand{\notimplementedyet}[1]{\color{blue}\emph{#1}\footnote{Still needs to be implemented}\color{black}\xspace}
\algrenewcommand{\algorithmiccomment}[1]{\# \textit{#1}} % so that comments in algorithms are preprended by '#' instead of a small right arrow

\title{Boutiques: a flexible framework for automated application integration in computing platforms}

\begin{document}

\author{Tristan Glatard, Tristan Aumentado-Armstrong, Natacha Beck, Pierre Bellec,\\
        Remi Bernard, Sorina Camarasu-Pop, Fr\'ed\'eric Cervenansky, Samir Das, \\
        Rafael Ferreira da Silva, Guillaume Flandin, John Flavin, Pascal Girard, \\
         Krzysztof J. Gorgolewski, Charles G. Guttmann, Nathaniel Kofalt, Pierre-Olivier Quirion,\\
         Pierre Rioux, Gunnar Schaeffer, Marc-\'Etienne Rousseau, Alan C. Evans}

\maketitle

\abstract{Porting applications to execution platforms such as web
  portals and workflow engines is critical to enable their sharing
  among scientific communities and their efficient execution on
  computing infrastructures. However, application porting remains a
  costly human effort that consists of 1)~installing the application
  on the target infrastructure, 2)~describing the application in a
  format compatible with the execution platform, and 3)~generating
  proper user interfaces. Due to the variety of platforms, application
  porting efforts are often replicated several times while
  mutualization would save cost and improve the quality of the ported
  applications. We describe \boutiques, a system that allows for
  automatic import and exchange of applications across
  platforms. \boutiques adopts a rich, flexible application
  description format and it relies on Linux containers to install
  applications across platforms in a lightweight manner. It is
  currently supported by \todo{X} platforms and was used to describe
  \todo{Y} applications.  }

\section{Introduction}

Platforms such as web services, portals, virtual research
environments, science gateways and workflow engines often rely on
third-party data processing and simulation applications to deliver
services to their users. Such applications are often repeatedly and
manually ported to various platforms whereas this porting effort would
greatly benefit from being mutualized. Meanwhile, virtual container
systems such as Docker and Singularity greatly facilitate the sharing
of software and improve the reproducibility of analyses by defining
immutable, reusable execution environments. However, containers need
to be properly annotated to allow platforms to leverage the software
they contain. Basically, the command lines that can be run in a
container must be formally described so that they can be automatically
created and executed by platforms. Describing command lines is not
straightforward since they often contain implicit semantics specified
in human-readable documents such as manuals.

\boutiques is a system to share command-line applications across
platforms. In \boutiques, a command line is described using a flexible
template comprising the inputs it requires and the outputs it
produces. Inputs may be inter-dependent, e.g., input A cannot be used
if input B is used, and passed in configuration files or directly on
the command line. Such formal descriptions, also called
\emph{manifests}, allow for advanced validation of input values to
prevent errors. Manifests refer to containers where applications and
all their dependencies are installed. \boutiques also provides a
simple parallelization framework that allows applications such as
workflow engines to concurrently execute multiple tasks (in the
remainder, tasks are called \emph{invocations}). A set of core tools
facilitate the construction, validation and execution of \boutiques
manifests. A repository orchestrates these tools together. \boutiques
manifests are intented to be produced by scientific application
developers, to be published in common repositories and to be consumed
by execution platforms.

The remainder of this paper describes the \boutiques system
(Section~\ref{sec:system}) and reports on its adoption by platforms
and applications in the neuroinformatics domain, our primary field of
interest (Section~\ref{sec:results}). It closes on a discussion and
comparison with related systems.

\note{Tristan}{Vocabulary (check consistency across the paper): tool vs. application}
\note{Tristan}{Targeted journals: Scientific Data, Giga Science, Frontiers in Neuroinformatics, Future Generation Computer Systems}

\section{System description}
\label{sec:system}

Applications are described with a JSON manifest that specifies the
command-line template, inputs and outputs. The manifest may point to a
container where the application and all its dependencies are
installed. It may also contain an invocation schema used for input
validation. At runtime, the execution platform builds the command line
from the manifest and values entered by the user. The platform runs
the command line on the execution infrastructure, e.g., a server, a
cluster or a cloud, within a container whenever available. To build
and run the command line, the platform may rely on the \boutiques core
tools, in particular on the validator and local executor. Application
manifests may be stored in the repository which provides additional
features.

\subsection{Command-line description}

The core component of the manifest is a flexible command-line
description represented as a simple string template complying to the
syntax of the \texttt{bash} UNIX shell. \texttt{bash} is the default
shell on the major Linux distributions and OS X. The command-line
template may contain placeholders for input and output values, called
value keys. The command line may actually
encompass several commands separated, e.g., by semicolumns, pipes
(\texttt{|}) or ampersands (\texttt{\&}). Such a flexible command-line
specification is meant to facilitate the embedding of minor mundane
operations on the command line, for instance directory creation, input
decompression or output archival.

Here is an example of a typical command-line template:
\begin{verbatim}
exampleTool_1 [STRING_INPUT] [FILE_INPUT] [ENUM_INPUT] | \
                              exampleTool_2 [FLAG_INPUT] [NUMBER_INPUT] >> [LOG].txt
\end{verbatim}
The template contains six value keys for input and output values, that
will be replaced by values and file names according to the user input
when the application is executed. Flags will also be added wherever
appropriate, with customizable separators. No particular syntax is
imposed for the value keys, they just have to be unique.  Note the use
of the pipe (\texttt{|}) operator to chain applications, and of the
\texttt{>>} operator to redirect the standard output to a file.

\subsection{Input description}

\paragraph{General properties.} Inputs must have a name, a unique
identifier and a type. They may be optional, have a description, a
value key, a flag and flag separator, and a default value. Inputs may
also be ordered lists: in this case, value keys are substituted by the
space-separated list of input values.

\paragraph{Types.} Inputs may be of type \texttt{String},
\texttt{Number}, \texttt{Flag} or \texttt{File}. \texttt{File} may
also represents a directory. Types can be restricted to a specific set
of values (\texttt{String} and \texttt{Number} only), to a specific
range (\texttt{Number} only), to integers or to booleans (\texttt{Number}
only).

\paragraph{Groups and dependencies.} Groups of inputs may be defined
with a identifier, name and list of input identifiers. Groups may be
used to improve the presentation in a graphical user interface and to
specify the following constraints among inputs: (1)
\texttt{mutually-exclusive}: only one input in the group may have a
value; (2) \texttt{one-is-required}: at least one input in the group
must have a value. Dependencies among inputs may also be defined
regardless of a particular group: an input may (1) require a list of
inputs and (2) disable a list of inputs.

Listing~\ref{listing:input-example} shows the definition of an input
in the command line exemplified above. According to this definition
and assuming that the input value entered by the user is 0.3, the
string \texttt{[NUMBER\_INPUT]} will be replaced by \texttt{-n=0.3} on
the command line at execution time.
\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=21pt,
               tabsize=4]{js}
{
    "id" : "num_input",
    "name" : "A number input",
    "type" : "Number",
    "value-key" : "[NUMBER_INPUT]",
    "optional" : true,
    "command-line-flag" : "-n",
    "command-line-separator" : "=",
    "minimum" : 0,
    "maximum" : 1,
    "exclusive-minimum" : true,
    "exclusive-maximum" : false
  }
\end{minted}
\caption{Input example.} 
\label{listing:input-example}
\end{listing}

\subsection{Output description}

Application outputs are files and directories that need to be
transferred and delivered to the user once the execution is
complete. Although it may not be required to distinguish inputs from
outputs to build the application command line, this information is
needed by computing platforms to identify the files that must be saved
after the execution.

In \boutiques, output files must have a unique identifier, a name and
a path template that specifies the file or directory name. Path
templates may include input value keys in case output files are named
after the input values. In this case, input values may be stripped
from specific strings, e.g., file extensions before being substituted
in the path template. Output files may also have a description, a
command-line flag, a flag separator, and a value key in case they
appear on the command line. They may be optional in case the file is
not always produced by the application, for instance when it is
produced only when a particular flag is activated. They may also be
(un-ordered) lists: in this case, the path template must contain a
wildcard (\texttt{*}) matching any string of characters and defining
the pattern used to match the output files in the list.

Listing~\ref{listing:output-example} shows the definition of an output
file in the command line exemplified before. According to this
definition and assuming that the string input value entered by the
user is "foo.csv", the string \texttt{[LOG]} will be
replaced by \texttt{log-foo} on the command line at execution time.

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=21pt,
               tabsize=4]{js}

{
    "id" : "logfile",
    "name" : "Log file",
    "description" : "The output log file from the example application",
    "path-template" : "log-[STRING_INPUT]",
    "value-key" : "[LOG]",
    "path-template-stripped-extensions" : [".txt", ".csv"],
    "optional" : false
}
\end{minted}
\caption{Output file example} 
\label{listing:output-example}
\end{listing}

\subsection{Configuration files}

A large number of applications rely on configuration files rather than
command-line options to define their input and output
parameters. Indeed, as the number of parameters increases, command
lines rapidly become long and cumbersome whereas configuration files
allow for better structure and documentation.

Configuration files may be complex though, and specified in any
language.  As it does not seem reasonable to adopt a particular
template for configuration files, the \boutiques specification allows
application developers to specify their own template containing input
and output value keys. Configuration files are specific types of
output files that must have a file template in addition to a path
template that defines how they will be named and where they will be
written. They may also have a value key and a flag in case they need
to be passed on the command
line. Listing~\ref{listing:configuration-file-example} shows an
example.
\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=21pt,
               tabsize=4]{js}
               {
                   "id": "config_file",
                   "name": "Configuration file",
                   "type": "Configuration File",
                   "value-key": "[CONFIG_FILE]",
                   "path-template": "config.txt",
                   "file-template": [
                   "# This input is hard-coded",
                   "stringInput=foo",
                   "fileInput=[FILE-INPUT]",
                   "# And here is the result",
                   "fileOutput=[OUTPUT-FILE]",
                   ""
                   ]
                 }
\end{minted}
\caption{Configuration file example. The file template is defined as
  an array of strings to allow for multi-line strings in JSON.}
\label{listing:configuration-file-example}
\end{listing}

\subsection{Command-line construction}

At runtime, a value is assigned to all the mandatory and some of the
optional inputs. %% Inputs of type ``String'' may contain any string,
%% inputs of type "Number" must contain a string representing a number,
%% inputs of type "File" must contain a string representing a file path
%% (absolute or relative to the execution directory) and inputs of type
%% "Flag" must contain a boolean. 
%% When input is a list, the value contains the concatenation of all the
%%values in the list.
Algorithm~\ref{algo:command-line} shows how the command line is
constructed from the manifest and the values entered by the user. It
substitutes all the value keys in the command line, output path
templates and configuration files, and writes the configuration files.
\begin{algorithm}[h!]
\caption{Command-line construction}
\label{algo:command-line}
\begin{algorithmic}
  \State \Comment{Substitute input value keys in output path templates, configuration files and command line.}
  \For{\texttt{input} in \texttt{inputs}}
  \If{\texttt{input} has a \texttt{value-key}}
  \For{\texttt{output} in \texttt{outputs}}
  \State \texttt{stripped\_value} = \texttt{input\_value}
  \If{\texttt{input} type is \texttt{File}}
  \State In \texttt{stripped\_value}, remove  all elements in \texttt{path-template-stripped-extensions}.
  \EndIf
  \State In \texttt{path-template}, replace all occurrences of \texttt{value-key} by \texttt{stripped\_value}.
  \If{\texttt{output} has a \texttt{file-template}}
  \State In any line of \texttt{file-template}, replace all occurrences of \texttt{value-key} by \texttt{stripped\_value}.
  \EndIf
  \EndFor
  \State Prepend \texttt{command-line-flag} and \texttt{command-line-flag-separator} to \texttt{input\_value}.
  \State In \texttt{command-line}, replace all occurrences of \texttt{value-key} by \texttt{input\_value}.
  \EndIf
  \EndFor

  \State \Comment{Substitute output value keys in configuration files and command line.}
  \For{\texttt{output} in \texttt{outputs}}
  \If{\texttt{output} has a \texttt{value-key}}
  \For{\texttt{output} in \texttt{outputs}}
  \If{\texttt{output} has a \texttt{file-template}}
  \State In any line of \texttt{file-template}, replace all occurrences of \texttt{value-key} by \texttt{path-template}.
  \State \Comment{Input \texttt{value-key} have been substituted in \texttt{path-template} previously.}
  \EndIf
  \EndFor
  \State Prepend \texttt{command-line-flag} and \texttt{command-line-flag-separator} to \texttt{path-template}.
  \State In \texttt{command-line}, replace all occurrences of \texttt{value-key} by \texttt{path-template}.
  \EndIf
  \EndFor

  \State \Comment{Write all configuration files.}
  \For{\texttt{output} in \texttt{outputs}}
  \If{\texttt{output} has a \texttt{file-template}}
  \State Write \texttt{file-template} in \texttt{path-template}
  \State \Comment{Value keys have been substituted in \texttt{file-template} previously.}
  \EndIf
  \EndFor

\end{algorithmic}
\end{algorithm}

\subsection{Invocation schema}
\label{sec:invocation-schema}

Rigorous input validation is an important motivation for
\boutiques. The obvious solution for JSON documents is to rely on an
application-specific JSON schema, called \emph{invocation schema}, to
specify the input values accepted by an application. Platforms can
rely on invocation schemas to validate inputs using
any JSON validator, without having to develop specific code.

Invocation schemas, however, are complex JSON objects. Basically, they
must represent the properties described above in a formal way,
including dependencies between
inputs. Listing~\ref{listing:invocation-schema-example} shows an
example of how dependencies between mutually exclusive parameters are
defined in the invocation schema. As it does not seem realistic to
assume that developers will easily write correct invocations schemas
even for their own applications, invocation schemas can be generated
automatically using a core \boutiques tool. The invocation schema is
stored as an optional property of the Boutiques manifest.

\notimplementedyet{The invocation schema also has a
  \texttt{task-dependencies} array that allows specifying that the new
  invocation may start only when previously-submitted invocations have
  completed. This property is used in the parallelization model
  described hereafter.}

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=21pt,
               tabsize=4]{js}
"dependencies" : {
      "num_input" : {
         "properties" : {
            "str_input" : {
               "not" : {}
            }
         }
      },
      "str_input" : {
         "properties" : {
            "num_input" : {
               "not" : {}
            }
         }
      }
   }
\end{minted}
\caption{Excerpt from invocation schema showing dependencies between
  two mutually exclusive parameters \texttt{num\_input} and
  \texttt{str\_input}.}
\label{listing:invocation-schema-example}
\end{listing}

\subsection{Parallelization support}
\label{sec:parallelization}
\boutiques supports parallelism by allowing applications to submit new
invocations through the following protocol \todo{Needs to be
  documented on Github}:
\begin{enumerate}
\item Applications write a JSON object complying to the invocation
  schema in their execution directory.
\item The platform reads the JSON object, creates the corresponding
  invocation, and writes back the invocation id in the same directory.
% The platform could also support status requests.
\end{enumerate}
Several invocations can be submitted at once, possibly using the
\texttt{task-dependencies} array to specify dependencies among
them. This mechanism is implemented in the Boutiques schema using a
single property, \texttt{can-submit-tasks}.

Although very simple, this model allows wrapping complete
workflows. Workflows are wrapped as any other application, and they
can be executed using any engine provided that it is installed in the
container image referred in the application manifest. Such a ``task
encapsulation'' model allows for a scalable and reliable execution of
workflows expressed in a variety of languages, as detailed in
~\cite{glatard2016fgcs}.

\subsection{Containers}

Applications are installed in a container image complying to the
Docker, Singularity or rootfs format. We intentionally support
multiple container formats as we anticipate that they will be used for
different purposes. For instance, Docker is well suited for developers
and users who manipulate applications on their local workstation. It
is well documented, maintained and it has a rich ecosystem of tools to
help build and run containers on most operating systems. Singularity
is more suited for users and platforms that need to run applications
on shared computing clusters. Bridges exist among these containers
formats to convert container images across frameworks. For instance, a
platform dedicated to high-performance computing may accept manifests
referring to Docker containers to facilitate application integration
by developers, and internally convert container images to Singularity
to run applications efficiently on clusters.

Container images are defined from their URL (Singularity and rootfs)
or name in a Docker index. Container images may specify a working
directory where the application has to be run. A hash may also be
reported to accurately identify container images and detect updates.

Containers were adopted because they allow for an automated and
lightweight integration of application implementations in
platforms. They are extremely useful to improve the reproducibility of
analyses as variations in the software environment may have an
important impact on the computed results. They also have limitations,
in particular they do not specify the hardware architecture required
to execute an application, which can be an issue in some cases.

Containers are not mandatory though. If an application manifest does
not contain any reference to a container, the platform will assume
that the application is pre-installed on the execution infrastructure.

\subsection{Custom properties}

Custom properties may be added to the Boutiques specification without
restriction. Custom properties are corralled in a specific JSON object
to facilitate validation. They may be useful to implement
platform-specific features but they should be used with care to avoid
making applications dependent on a particular platform.

\subsection{Core tools}

The Boutiques validator checks conformance of JSON manifests to the
Boutiques schema using a basic JSON validator. It also performs the
following checks that cannot be easily implemented in JSON schema:
value keys are unique among inputs, input and output identifiers are
unique, input and output value keys are all included in the command
line, value keys are not contained within each other (which would
puzzle substitution), output path templates are unique (to avoid
results to overwrite each other), inputs of type Flag have a
command-line flag, are optional and are not lists, number and list
constraints are sensible (e.g. min is lower than max), the default
value of enum types is part of the enum, an input cannot both require
and disable another input, required inputs cannot require or disable
other parameters, group member identifiers must correspond to existing
inputs and cannot appear in different groups, mutually exclusive
groups cannot have members requiring other members, and
one-is-required groups should never have required members. The
Boutiques validator is meant for application developers; platforms
should rely on regular JSON validation to avoid depending on a
third-party tool.

The local executor can generate and execute command line from a
Boutiques manifest and a set of input values represented in a CSV or
JSON file complying with the invocation schema. It runs the command in
a container provided that the required framework (e.g. Docker) is
already installed. It can also generate hypothetical command lines
from random values for debugging purposes. The local executor can be
used by application users to run applications locally, or by platforms
to generate command lines to be run on the execution infrastructure.

The invocation schema generator can create an invocation schema (see
Section~\ref{sec:invocation-schema}) from a \boutiques manifest and
validate input data against it using a regular JSON validator. It can
be used to add invocation schemas to existing manifests.

Finally, a JSON editor facilitates the writing of \boutiques manifests
through a web interface.

\subsection{Repository}

\todo{Gunnar, Nathaniel, John}

\section{Results}
\label{sec:results}

\subsection{Supported platforms}

The following platforms are able to integrate and execute Boutiques
applications.

\subsubsection{CBRAIN}

CBRAIN~\cite{SHER-14} is a web platform that can process data
distributed into multiple storage locations on computing clusters and
clouds. CBRAIN offers transparent access to remote data sources,
distributed computing sites, and an array of processing and
visualization tools within a controlled, secure environment.  The
CBRAIN service deployed at the Montreal Neurological Institute relies
on the infrastructure provided by Compute Canada~\cite{das2016mni}. It
currently provides 460+ collaborators in 20 countries with web access
to several systems, including 6 clusters of the Compute Canada
high-performance computing infrastructure (totalling more than 100,000
computing cores and 40PB of disk storage) and Amazon EC2. CBRAIN
transiently stores about 10 million files representing over 50TB
distributed in 42 servers. 56 data processing applications are integrated and
over 340,000 processing batches have been submitted since 2010.

Applications in CBRAIN are integrated as Ruby classes that create web
forms, validation and execution code. The Boutiques support in CBRAIN
is based on a set of templates that generate such classes from the
application manifest. Two integration modes are available to integrate
Boutiques applications: (1) the application manifest is stored in a
CBRAIN plugin, and the Ruby classes are generated on-the-fly when
CBRAIN starts. This mode allows CBRAIN developers to add new features
to all Boutiques applications (e.g. improved error message reporting)
by updating the Ruby templates. It is, however, constrained to the
Boutiques manifest schema: features not supported in Boutiques, e.g.,
cluster-specific adjustments, cannot be supported. (2) CBRAIN Ruby
classes are generated from the Boutiques manifest in an offline
process and integrated in CBRAIN as any other application. This mode
allows developers to customize applications by editing the generated
Ruby classes, but the resulting applications are difficult to maintain
in the long term, in particular when the manifests are updated.

A custom property (\texttt{cbrain:inherits-from-class}) was added to
the Boutiques manifest to define the Ruby class that should be used as
parent class for the application in CBRAIN. It provides more
flexibility to the first integration model described previously. For
instance, it was used to define a progress bar for PSOM pipelines in
CBRAIN.

CBRAIN was also extended to support the parallelization mechanism
available in Boutiques: the CBRAIN execution agent periodically scans
the execution directories of the parallelized applications and it submits
the invocations defined in the JSON objects written by the
application in this directory.

A new list mechanism was also introduced in CBRAIN to facilitate the
iteration of Boutiques applications on large sets of files. CBRAIN
lists are specific files that contain references to other CBRAIN
files. When a list is passed to a Boutiques application, the elements
in the list are either concatenated in a single command line (when the
corresponding Boutiques input is a list), or a new command line is
generated for every element in the list (when the input is not a
list). Supporting lists as a specific CBRAIN file type allows for
improved validation, e.g. lists that contain references to
non-existent or deleted files can be detected. It also allows users to
edit lists using their own tools such as scripts or spreadsheet
applications.

\subsection{Flywheel}

\todo{Gunnar, Nathanael, John}

\subsection{Pegasus}

\todo{Rafael}

\subsection{SPINE}

\todo{Charles}

\subsection{VIP}

\subsection{Ported applications}

% cbrain-plugins-neuro

Table~\ref{table:applications} summarizes the applications described
with Boutiques and the main features used.

\paragraph{Niak.} The Niak pre-processing pipeline, executed with the Pipeline System
for Octave and Matlab (PSOM)~\cite{bellec2012pipeline}, was integrated
in CBRAIN using Boutiques. The integration uses the parallelization
mechanism described in Section~\ref{sec:parallelization} so that even
the invocations processing a single subject can be parallelized. It also
allows CBRAIN to leverage the efficient agent model used in PSOM, as
described in~\cite{GLAT-16}. The integration required some work in
PSOM to facilitate its invocation as a non-interactive command-line
application. The resulting CBRAIN plugin is available at
\url{https://github.com/SIMEXP/cbrain-plugins-psom}. It could in
principle be used in other platforms that support the parallelization
property, although this has not been tested.

\paragraph{Nipype.} The Nipype workflow engine~\cite{gorgolewski2011nipype} was
instrumented to export to Boutiques applications described in its
internal Python description language. \todo{Update nipype\_cmd to
  support the latest Boutiques schema and complete this paragraph.}

\paragraph{SPM.} The Statistical Parameter Mapping toolbox (SPM)~\cite{penny2011statistical}
was exported to Boutiques through a specific \texttt{spm2boutiques}
tool that writes Boutiques manifests from configuration files of SPM
functions (object declarations). \todo{More on that when it is
  completed.}

\paragraph{MICCAI challenges.} Boutiques was also used to port to the VIP platform the 23 pipelines
involved in two computational challenges organized by the MICCAI
conference in 2016, related to the segmentation of multiple-sclerosis
lesions in MR images and of tumor volumes in PET images. The pipeline
developers created Boutiques manifests and wrapped their applications
in Docker containers. The pipelines were ported to VIP and executed on
205 subjects in a few weeks only. Some pipelines had to be adjusted
once ported to the platform for the following reasons:
\begin{enumerate}
\item A pipeline
  required a GPU, which we enabled through the
  \texttt{nvidia-docker}\footnote{\url{https://github.com/NVIDIA/nvidia-docker}}
  tool not supported in Boutiques by default although it could be a possible extension.
\item A pipeline required
  more than 10 GB of data dependencies (atlas data) which exceeded the
  maximum size allowed for Docker containers in our setup; we solved the
  issue by installing the data on a directory of the host server that we
  mounted in the container. This could not possibly be supported by
  Boutiques since it is very dependent on the platform.
\item A pipeline wrote more than 10 GB of intermediate data in a
  temporary directory of the container located on a 2 GB partition. We
  solved the issue by mounting a host directory in the temporary directory.
\end{enumerate}
The pipelines did not use much of the expressiveness features offered
by Boutiques since the instructions given to the challengers required
that all the pipeline parameters were fixed, to facilitate
execution. Two custom properties (\texttt{vip:miccai-challenger-email}
and \texttt{vip:miccai-challenge-team-name}) were added to the
Boutiques manifest to help VIP handle pipeline results for the 2016
MICCAI challenges.

\paragraph{Other applications.} Other neuroinformatics applications were ported to CBRAIN using Boutiques:
FSL tools and ICA AROMA are available at
\url{https://github.com/aces/cbrain-plugins-neuro} and some pipelines
of the Human Connectome Project (HCP) are at
\url{https://github.com/big-data-lab-team/cbrain-plugins-hcp}.

\todo{Chris, something on BIDS apps?}

\todo{Sorina, a paragraph on other applications in VIP?}

\begin{table}
  \begin{tabular}{l|ccccccc}
    \rowcolor[gray]{0.9}
    Application                      & \multicolumn{7}{c}{Input/Output properties} \\
    \rowcolor[gray]{0.9}
                                     & Dependencies      &Groups             & Lists             & Optional          &Default            & Enum              & Min/max\\

    \hline
    Niak pre-processing              &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\\
    ICA AROMA                        &\cellcolor{gray!75}&\cellcolor{gray!75}&                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
    FSL tools &&&&&&\\
    \multicolumn{1}{r|}{anat}        &\cellcolor{gray!75}&\cellcolor{gray!75}&                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
    \multicolumn{1}{r|}{bet}         &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&                   &                   &\cellcolor{gray!75}\\      
    \multicolumn{1}{r|}{fast}        &\cellcolor{gray!75}&                   &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
    \multicolumn{1}{r|}{first}       &                   &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&                   &                   \\
    FSL Nipype interfaces     \\
    HCP PreFreesurfer                &                   &                   &                   &                   &                   &                   &                   \\
    MICCAI challenge (23 pipelines)\\&                   &                   &                   &                   &                   &                   & \\
    \end{tabular}
  \vspace*{0.5cm}
  \begin{tabular}{l|ccc|ccc}
    \hline
    \rowcolor[gray]{0.9}
    Application                     &\multicolumn{3}{c|}{Containers}            &  \multicolumn{3}{c}{Other properties}  \\
    \rowcolor[gray]{0.9}
                                    &Docker              & Singularity& Rootfs  & Config. file & Env vars  & Parallelization   \\
    \hline
    Niak pre-processing             &\cellcolor{gray!75} &            &         &              &\cellcolor{gray!75}&\cellcolor{gray!75}\\
    ICA AROMA                       &\cellcolor{gray!75} &            &         &              &&\\
    FSL tools &&&&&\\
    \multicolumn{1}{r|}{anat}       &                    &            &         &              &&\\
    \multicolumn{1}{r|}{bet}        &                    &            &         &              &&\\
    \multicolumn{1}{r|}{fast}       &                    &            &         &              &&\\
    \multicolumn{1}{r|}{first}      &                    &            &         &              &&\\
    FSL Nipype interfaces     \\
    HCP PreFreesurfer               &\cellcolor{gray!75} &            &         &             &&\\
    MICCAI challenge (23 pipelines) &\cellcolor{gray!75} &            &         &             &&\\
  \end{tabular} 
  \caption{Applications available in Boutiques with manifest features used.}
  \label{table:applications}
\end{table}

\section{Discussion}

\begin{itemize}
\item Security: containers may be updated after being integrated in the platforms.
\item The manifest specification is getting increasingly complex, but
  this complexity is required to describe real-world
  applications. When something cannot be described using the manifest,
  it is always possible to change the command line format by writing a
  wrapper script for the application. However, when containers are
  used, the wrapper script needs to be added to the container, which
  has some overhead (in particular, the containers provided by the
  developers cannot be used directly).
\item Testing is critical.
\item Workflow engines can be platforms (they can run Boutiques applications), workflows can be applications (they can be described in Boutiques).
\item With Boutiques, developers can port their application once and execute it in several platforms. It removes the technological dependency to a particular platform and facilitates migration. 
\end{itemize}

The \boutiques specification adopts a bottom-up approach where new
features are progressively added when applications or platforms
require them. It is currently supported by \todo{X} platforms and  \todo{Y} applications
were described.

 Although the motivating use cases
were taken from neuroinformatics, our primary field of interest,
nothing prevents the system from being used in a variety of other
fields.

\section{Related work}

Boutiques intentionally does not support the definition of complete
workflows from individual applications. Indeed, we believe that
workflow engines are specific applications that need complex languages to
implement the particular functions that they target. Efforts to
specify common workflow languages include the Common Workflow Language
(CWL)~\cite{cwl} and the Interoperable Workflow Intermediate
Representation (IWIR)~\cite{plankensteiner2011iwir}.

% CWL
The Common Workflow Language (CWL) is the closest related work to
Boutiques, in the sense that it also allows the description of command
lines executed in containers. According to Github, the CWL project
started 6 months before Boutiques. The main conceptual differences
between CWL and Boutiques are the following:
\begin{itemize}
\item CWL has a workflow language, Boutiques does not. CWL also has a
  command line tool description language, but its specification is
  influenced by the fact that command line tools are included in a
  specific workflow languages. In Boutiques, the only available
  workflow construct is sub-tasking (``a task may submit other
  tasks''). This is intentional as we envisage that Boutiques tools
  may be composed through any type of workflow engine and language. For instance, this leads to have multiple output types in CWL.
\item CWL specifies command lines formally whereas in Boutiques they
  are assembled from simple string templates. In short, a command line
  in CWL is specified by an executable and an ordered list of
  parameters. While this representation has the advantage of being
  more formal, for instance the executable can be identified
  automatically, it also removes flexibility by imposing that a tool
  only runs a single command. On the contrary, Boutiques allows any
  type of bash operators to be combined on the command line. This
  level of flexibility is important and greatly reduces the ``shim
  syndrome''. Boutiques' template approach allows for configuration files. 
\item CWL is validated using SALAD, which makes the ecosystem slightly
  more complex. For instance, CWL has ``import'' and ``include''
  directives. Boutiques is validated using plain JSON schema. \todo{ontology layer}.
\item In CWL, the applications need to be instrumented as follows, which is not the case in Boutiques:
\begin{itemize}
\item Applications must write in specific directories.
\item Applications must produce a JSON object describing the output structure.
\end{itemize}
\end{itemize}

\todo{Update from comparison summary on Google doc}

% How to build modular container images for workflows.

\section{Conclusion}

\section{Acknowledgments}

Amazon grant.

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}