### 2017-10-19 boutiquespaper.tex

\documentclass{article}

%%%% Packages %%%%

\usepackage{minted} % Used for JSON highlighting
\usepackage{pdfcomment} % Used for notes and todos
\usepackage[margin=1in]{geometry} % Increases margins
\usepackage[table]{xcolor} % Color definitions
\usepackage{algorithm} % Algorithm float
\usepackage{algpseudocode} % Algorithmic environment
\usepackage{xspace}

%%%% Commands %%%%

\newcommand{\note}[2]{\pdfmargincomment[color=yellow,author=#1,open=true]{#2}}
\newcommand{\todo}[1]{\color{red}TODO: #1\color{black}}
\newcommand{\boutiques}{Boutiques\xspace}
\newcommand{\notimplementedyet}[1]{\color{blue}\emph{#1}\footnote{Still needs to be implemented}\color{black}\xspace}
\algrenewcommand{\algorithmiccomment}[1]{\# \textit{#1}} % so that comments in algorithms are preprended by '#' instead of a small right arrow

\title{Boutiques: a flexible framework for automated application integration in computing platforms}

\begin{document}

\author{Tristan Glatard, Gregory Kiar, Tristan Aumentado-Armstrong,
  Natacha Beck, Pierre Bellec,\\ Remi Bernard, Sorina Camarasu-Pop,
  Fr\'ed\'eric Cervenansky, Samir Das, \\ Rafael Ferreira da Silva,
  Guillaume Flandin, John Flavin, Pascal Girard, \\ Krzysztof
  J. Gorgolewski, Charles G. Guttmann, Val\'erie Hayot-Sasson,
  Nathaniel Kofalt, \\ Pierre-Olivier Quirion, Pierre Rioux,
  Marc-\'Etienne Rousseau, Gunnar Schaeffer, Alan C. Evans}

\maketitle

\abstract{Porting applications to execution platforms such as web
  portals and workflow engines is critical to enable their sharing
  among scientific communities and their efficient execution on
  computing infrastructures. However, application porting remains a
  costly human effort that consists of 1)~installing the application
  on the target infrastructure, 2)~describing the application in a
  format compatible with the execution platform, and 3)~generating
  proper user interfaces. Due to the variety of computing platforms,
  application porting efforts are often replicated several times while
  interoperability would save cost and improve the quality of the ported
  applications. We describe \boutiques, a descriptive framework and
  command line utility that allows for automatic import and exchange of
  applications across platforms. \boutiques adopts a rich, flexible
  application description format which relies on Linux container
  environments such as Docker and Singularity to install applications
  across platforms in a lightweight manner. \boutiques is currently used
  by three distinct virtual research platforms, and has been used to
  describe dozens of neuroinformatics tools.}

\section{Introduction}

Platforms such as web services, portals, and virtual research
environments are becoming an common solution for increasing the
accessibility of high performance computing. These scientific gateways and
workflow engines often rely on third-party data processing and simulation
applications to deliver services to their users. Such applications are
often repeatedly and manually ported to various platforms whereas this
porting effort would greatly benefit from being mutualized. Meanwhile,
virtual container systems such as Docker and Singularity greatly facilitate
the sharing of software and improve the reproducibility of analyses by defining
immutable, reusable execution environments. However, containers need
to be properly annotated to allow platforms to leverage the software
they contain. The command lines that can be run in a container must be formally
described so that they can be automatically created and executed by platforms.
Describing command lines is not straightforward since they often contain
implicit semantics specified in human-readable documents such as manuals.

\boutiques is a system to share command-line applications across
platforms. In \boutiques, a command line is described using a flexible
template comprising the inputs it requires and the outputs it
produces. Inputs may be inter-dependent, e.g., input \emph{A} cannot be used
if input \emph{B} is used; inputs may be passed in configuration files
or directly on the command line. Such formal descriptions, simply referred to
as \emph{descriptors}, allow for advanced validation of input values to
prevent errors. A descriptor refers to a container where a given application
and all their dependencies are installed. \boutiques also provides a
simple parallelization framework that allows applications such as
workflow engines to concurrently execute multiple tasks (in the
remainder, tasks are called \emph{invocations}). A set of core tools
facilitate the construction, validation, import, execution, and publishing
of \boutiques descriptors. \boutiques descriptors are intented to be produced by
scientific developers and stored alongside their application, indexed by common
repositories, and to be consumed by execution platforms.

The remainder of this paper describes the \boutiques system
(Section~\ref{sec:system}) and reports on its adoption by platforms
and applications in the neuroinformatics domain, our primary field of
interest (Section~\ref{sec:results}). It closes on a discussion and
comparison with related systems.

%\note{Tristan}{Vocabulary (check consistency across the paper): tool vs. application}
%\note{Tristan}{Targeted journals: Scientific Data, Giga Science, Frontiers in Neuroinformatics, Future Generation Computer Systems}

\section{System description}
\label{sec:system}

Applications are described with a JSON descriptor that specifies the
command-line template, inputs and outputs. The descriptor may point to a
container where the application and all its dependencies are
installed. It may also contain an invocation schema used for input
validation (this will be created at runtime if it is not found). At
runtime, the execution platform builds the command line from the descriptor
and values entered by the user. The platform runs the command line on the
execution infrastructure, e.g., a server, a cluster or a cloud, within a
container whenever available. To build and run the command line, the
platform may rely on the \boutiques core tools, packaged through the 
\texttt{bosh} command-line utility, such as the validator and executor.

\subsection{Command-line description}

The core component of the descriptor is a flexible command-line
description represented as a simple string template complying to the
syntax of the \texttt{bash} UNIX shell. \texttt{bash} is the default
shell on the major Linux distributions and OS X. The command-line
template may contain placeholders for input and output values, called
value keys. The command line may actually encompass several commands
separated by semicolons, pipes (\texttt{|}) or ampersands (\texttt{\&}).
Such a flexible command-line specification is meant to facilitate the
embedding of minor mundane operations on the command line, for instance
directory creation, input decompression or output archival.

Here is an example of a typical command-line template:
\begin{verbatim}
          exampleTool_1 [STRING_INPUT] [FILE_INPUT]  | \
                exampleTool_2 [FLAG_INPUT] [NUMBER_INPUT] >> [LOG].txt
\end{verbatim}
The template contains six value keys for input and output values that
will be replaced by values and file names according to the user input
when the application is executed. Flags will also be added wherever
appropriate, with customizable separators. No particular syntax is
imposed for the value keys, they just have to be unique.  Note the use
of the pipe (\texttt{|}) operator to chain applications, and of the
\texttt{>>} operator to redirect the standard output to a file.

\subsection{Input description}

\paragraph{General properties.} Inputs must have a name, a unique
identifier, and a type. They may be optional, have a description, a
value key, a flag and flag separator, and a default value. Inputs may
also be ordered lists: in this case, value keys are substituted by the
space-separated list of input values.

\paragraph{Types.} Inputs may be of type \texttt{String}, \texttt{Number},
\texttt{Flag} or \texttt{File}. \texttt{File} may also represent a directory.
Types can be restricted to a specific set of values (\texttt{String} 
and \texttt{Number} only), to a specific range (\texttt{Number} only), to
integers or to booleans (\texttt{Number} only).

\paragraph{Groups and dependencies.} Groups of inputs may be defined
with an identifier, name, and list of input identifiers. Groups may be
used to improve the presentation in a graphical user interface and to
specify the following constraints among inputs:
(1)~\texttt{mutually-exclusive}: only one member in the group may have a
value; (2)~\texttt{one-is-required}: at least one member in the group
must have a value; (3)~\texttt{all-or-none}: if any of the members are
present all members must be present. Dependencies among inputs may also be
defined regardless of a particular group: an input may (1)~require a list of
inputs and (2)~disable a list of inputs.

Listing~\ref{listing:input-example} shows the definition of an input in the
command line exemplified above. According to this definition and assuming
that the input value entered by the user is 0.3, the string
\texttt{[NUMBER\_INPUT]} will be replaced by \texttt{-n=0.3} on the command
line at execution time.
\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=21pt,
               tabsize=4]{js}
{
    "id" : "num_input",
    "name" : "A number input",
    "type" : "Number",
    "value-key" : "[NUMBER_INPUT]",
    "optional" : true,
    "command-line-flag" : "-n",
    "command-line-separator" : "=",
    "minimum" : 0,
    "maximum" : 1,
    "exclusive-minimum" : true,
    "exclusive-maximum" : false
}
\end{minted}
\caption{Example of a \texttt{Number}-type input.} 
\label{listing:input-example}
\end{listing}

\subsection{Output description}

Application outputs are files and directories that need to be
transferred and delivered to the user once the execution is
complete. Although it may not be required to distinguish inputs from
outputs to build the application command line, this information is
needed by computing platforms to identify the files that must be saved
after the execution.

In \boutiques, output files must have a unique identifier, a name, and
a path template that specifies the file or directory name. Path
templates may include input value keys in case output files are named
after the input values. In this case, input values may be stripped
from specific strings, e.g., file extensions, before being substituted
in the path template. Output files may also have a description, a
command-line flag, a flag separator, and a value key in case they
appear on the command line. They may be optional in case the file is
not always produced by the application, for instance when it is
produced only when a particular flag is activated. They may also be
(un-ordered) lists: in this case, the path template must contain a
wildcard (\texttt{*}) matching any string of characters and defining
the pattern used to match the output files in the list.

Listing~\ref{listing:output-example} shows the definition of an output
file in the command line exemplified before. According to this
definition and assuming that the string input value entered by the
user is "foo.csv", the string \texttt{[LOG]} will be
replaced by \texttt{log-foo} on the command line at execution time.

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=21pt,
               tabsize=4]{js}

{
    "id" : "logfile",
    "name" : "Log file",
    "description" : "The output log file from the example application",
    "path-template" : "log-[STRING_INPUT]",
    "value-key" : "[LOG]",
    "path-template-stripped-extensions" : [".txt", ".csv"],
    "optional" : false
}
\end{minted}
\caption{Example of an output leveraging \texttt{path-template} search-and-replacement.} 
\label{listing:output-example}
\end{listing}

\subsection{Configuration files}
\label{sec:configuration-files}
A large number of applications rely on configuration files rather than
command-line options to define their input and output
parameters. Indeed, as the number of parameters increases, command
lines rapidly become long and cumbersome whereas configuration files
allow for better structure and documentation.

Configuration files may be complex though, and specified in any
language.  As it does not seem reasonable to adopt a particular
template for configuration files, the \boutiques specification allows
application developers to specify their own template containing input
and output value keys. Configuration files are specific types of
output files that must have a file template in addition to a path
template that defines how they will be named and where they will be
written. They may also have a value key and a flag in case they need
to be passed on the command
line. Listing~\ref{listing:configuration-file-example} shows an
example.
\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=21pt,
               tabsize=4]{js}
{
    "id": "config_file",
    "name": "Configuration file",
    "type": "Configuration File",
    "value-key": "[CONFIG_FILE]",
    "path-template": "config.txt",
    "file-template": [
        "# This input is hard-coded",
        "stringInput=foo",
        "fileInput=[FILE-INPUT]",
        "# And here is the result",
        "fileOutput=[OUTPUT-FILE]",
        ""
    ]
}
\end{minted}
\caption{Example of a configuration input file. The file template is defined as
  an array of strings to allow for multi-line strings in JSON.}
\label{listing:configuration-file-example}
\end{listing}

\subsection{Command-line construction}

At runtime, a value is assigned to all the mandatory and some of the
optional inputs. %% Inputs of type ``String'' may contain any string,
%% inputs of type "Number" must contain a string representing a number,
%% inputs of type "File" must contain a string representing a file path
%% (absolute or relative to the execution directory) and inputs of type
%% "Flag" must contain a boolean. 
%% When input is a list, the value contains the concatenation of all the
%%values in the list.
%% gk: I agree with not including the above
Algorithm~\ref{algo:command-line} shows how the command line is
constructed from the descriptor and the values entered by the user. It
substitutes all the value keys in the command line, output path
templates and configuration files, and writes the configuration files.
\begin{algorithm}[h!]
\caption{Command-line construction}
\label{algo:command-line}
\begin{algorithmic}
  \State \Comment{Substitute input value keys in output path templates, configuration files and command line.}
  \For{\texttt{input} in \texttt{inputs}}
  \If{\texttt{input} has a \texttt{value-key}}
  \For{\texttt{output} in \texttt{outputs}}
  \State \texttt{stripped\_value} = \texttt{input\_value}
  \If{\texttt{input} type is \texttt{File}}
  \State In \texttt{stripped\_value}, remove  all elements in \texttt{path-template-stripped-extensions}.
  \EndIf
  \State In \texttt{path-template}, replace all occurrences of \texttt{value-key} by \texttt{stripped\_value}.
  \If{\texttt{output} has a \texttt{file-template}}
  \State In any line of \texttt{file-template}, replace all occurrences of \texttt{value-key} by \texttt{stripped\_value}.
  \EndIf
  \EndFor
  \State Prepend \texttt{command-line-flag} and \texttt{command-line-flag-separator} to \texttt{input\_value}.
  \State In \texttt{command-line}, replace all occurrences of \texttt{value-key} by \texttt{input\_value}.
  \EndIf
  \EndFor

  \State \Comment{Substitute output value keys in configuration files and command line.}
  \For{\texttt{output} in \texttt{outputs}}
  \If{\texttt{output} has a \texttt{value-key}}
  \For{\texttt{output} in \texttt{outputs}}
  \If{\texttt{output} has a \texttt{file-template}}
  \State In any line of \texttt{file-template}, replace all occurrences of \texttt{value-key} by \texttt{path-template}.
  \State \Comment{Input \texttt{value-key} have been substituted in \texttt{path-template} previously.}
  \EndIf
  \EndFor
  \State Prepend \texttt{command-line-flag} and \texttt{command-line-flag-separator} to \texttt{path-template}.
  \State In \texttt{command-line}, replace all occurrences of \texttt{value-key} by \texttt{path-template}.
  \EndIf
  \EndFor

  \State \Comment{Write all configuration files.}
  \For{\texttt{output} in \texttt{outputs}}
  \If{\texttt{output} has a \texttt{file-template}}
  \State Write \texttt{file-template} in \texttt{path-template}
  \State \Comment{Value keys have been substituted in \texttt{file-template} previously.}
  \EndIf
  \EndFor

\end{algorithmic}
\end{algorithm}

\subsection{Invocation schema}
\label{sec:invocation-schema}

Rigorous input validation is an important motivation for
\boutiques. The obvious solution for JSON documents is to rely on an
application-specific JSON schema, called \emph{invocation schema}, to
specify the input values accepted by an application. Platforms can
rely on invocation schemas to validate inputs using
any JSON validator, without having to develop specific code.

Invocation schemas, however, are complex JSON objects. Basically, they
must represent the properties described above in a formal way,
including dependencies between
inputs. Listing~\ref{listing:invocation-schema-example} shows an
example of how dependencies between mutually exclusive parameters are
defined in the invocation schema. As it does not seem realistic to
assume that developers will easily write correct invocations schemas
even for their own applications, invocation schemas can be generated
automatically by the \boutiques command-line utility. The invocation
schema is stored as an optional property of the \boutiques descriptor.

\notimplementedyet{The invocation schema also has a
  \texttt{task-dependencies} array that allows specifying that the new
  invocation may start only when previously-submitted invocations have
  completed. This property is used in the parallelization model
  described hereafter.}

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=21pt,
               tabsize=4]{js}
{
  "dependencies" : {
      "num_input" : {
         "properties" : {
            "str_input" : {
               "not" : {}
            }
         }
      },
      "str_input" : {
         "properties" : {
            "num_input" : {
               "not" : {}
            }
         }
      }
   }
}
\end{minted}
\caption{Excerpt from invocation schema showing dependencies between
  two mutually exclusive parameters \texttt{num\_input} and
  \texttt{str\_input}.}
\label{listing:invocation-schema-example}
\end{listing}

\subsection{Parallelization support}
\label{sec:parallelization}
\boutiques supports parallelism by allowing applications to submit new
invocations through \notimplementedyet{the following protocol}:
\begin{enumerate}
\item Applications write a JSON object complying to the invocation
  schema in their execution directory.
\item The platform reads the JSON object, creates the corresponding
  invocation, and writes back the invocation id in the same directory.
% The platform could also support status requests.
\end{enumerate}
Several invocations can be submitted at once, possibly using the
\texttt{task-dependencies} array to specify dependencies among
them. This mechanism is implemented in the Boutiques schema using a
single property, \texttt{can-submit-tasks}.

Although very simple, this model allows wrapping complete
workflows. Workflows are wrapped as any other application, and they
can be executed using any engine provided that it is installed in the
container image referred in the application descriptor. Such a ``task
encapsulation'' model allows for a scalable and reliable execution of
workflows expressed in a variety of languages, as detailed in
~\cite{glatard2016fgcs}.

\subsection{Containers}

Applications are installed in a container image complying to the
Docker, Singularity or rootfs format. We intentionally support
multiple container formats as we anticipate that they will be used for
different purposes. For instance, Docker is well suited for developers
and users who manipulate applications on their local workstations or the
cloud. It is well documented, maintained and it has a rich ecosystem of
tools to help build and run containers on most operating systems. Singularity
is more suited for users and platforms that need to run applications
on shared computing clusters. Bridges exist among these containers
formats to convert container images across frameworks. For instance, a
platform dedicated to high-performance computing may accept descriptors
referring to Docker containers to facilitate application integration
by developers, and internally convert container images to Singularity
to run applications efficiently on clusters.

Container images are defined from their URL (rootfs) or name in the
Docker or Singularity indices. Manifests and container images may specify a
working directory where the application has to be run, and a hash may also be
reported to accurately identify container images and detect updates.

Containers were adopted because they allow for an automated and
lightweight integration of application implementations in
platforms. They are extremely useful to improve the reproducibility of
analyses as variations in the software environment may have an
important impact on the computed results. They also have limitations,
in particular they do not specify the hardware architecture required
to execute an application, which can be an issue in some cases.

%%Containers are not mandatory though. If an application descriptor does
%%not contain any reference to a container, the platform will assume
%%that the application is pre-installed on the execution infrastructure. 
%% gk: I believe we only support containers now, correct?

\subsection{Custom properties}

Custom properties may be added to the Boutiques specification without
restriction. Custom properties are corralled in a specific JSON object
to facilitate validation. They may be useful to implement
platform-specific features but they should be used with care to avoid
making applications dependent on a particular platform or replacing
existing functionality already represented in the \boutiques schema.

\subsection{Core tools}

\boutiques is available on Python through the PyPi package repository as
``\texttt{boutiques}''. The \boutiques package exposes a command-line
utility to the user, \texttt{bosh}, which contains entrypoints for all
core functions within \boutiques. There is help-text which should guide
users through interacting with the \texttt{bosh} utility. The core tools
provided by \boutiques are, the: validator, executor (for both launch and
simulation), invocation schema handler, importer, and publisher. The tools
exposed through the command-line API are also available consistently through
the Python interface by importing the \boutiques package.

\paragraph{Validator.} The \boutiques validator checks conformance of JSON
descriptors to the \boutiques schema using a basic JSON validator. It also
performs the following checks that cannot be easily implemented in JSON schema:
value keys are unique among inputs, input and output identifiers are
unique, input and output value keys are all included in the command
line, identifiers with the same value key are mutually exclusive, value keys
are not contained within each other (which would puzzle substitution), output
path templates are unique (to avoid results to overwrite each other), inputs of
type Flag have a command-line flag, are optional and are not lists, the default
value of restricted types is part of the restriction, an input cannot both require
and disable another input, required inputs cannot require or disable
other parameters, group member identifiers must correspond to existing
inputs and cannot appear in different groups, mutually exclusive
groups cannot have members requiring other members, one-is-required groups
should never have required members, and all-or-none group members must not be
required.

\paragraph{Executor.} The executor has two modes of operation: \emph{launch}, and
\emph{simulate}. The simulate mode can generate hypothetical command lines from
random values given the descriptor (and corresponding invocation schema) for debugging
purposes, or display the command that would be executed given a provided valid
invocation. The launch mode can execute command line from a \boutiques descriptor and
a set of input values represented in JSON file complying with the invocation schema.
It runs the command in a container provided that the required framework (e.g. Docker)
is already installed. The executor can be used by application users to run
applications locally, or by platforms to generate command lines to be run on the
execution infrastructure.

\paragraph{Invocation Schema Handler.} The invocation schema handler can create an
invocation schema (see Section~\ref{sec:invocation-schema}) from a \boutiques descriptor
and validate input data against it using a regular JSON validator. It can be used to add
invocation schemas to existing descriptors. It is used by the executor if no invocation
schema is present in the \boutiques descriptor being deployed.

\paragraph{Importer.} The importer takes \boutiques descriptors from older versions and
updates them to be compliant with the most recent version of the schema. This tool can
also create descriptors from select application collections, such as BIDS
apps~\cite{gorgolewski2017bids}.

\paragraph{Publisher.} As \boutiques has primarily been adopted in the neuroinformatics
community, the publisher gets a further description (such as author, website, etc.) of
the described tool, and adds an index to it on
NeuroLinks\footnote{\url{https://brainhack101.github.io/neurolinks/}}, a repository
containing links to useful resources and tools. This enables others to find and use the
descriptor for your tool easily. This functionality could be extended to
new repositories for other domains, such as Bioconductor for bioinformatics.

Though not a component of the \boutiques tool chain, a Jupyter Notebook tutorial exists
to facilitate new users getting started with \boutiques.

\section{Results}
\label{sec:results}

\subsection{Supported platforms}

\boutiques application import and execution is currently supported by several
platforms which aid in execution and tool management. These are enumerated
below.

\subsubsection{CBRAIN}

CBRAIN~\cite{SHER-14} is a web platform that can process data
distributed into multiple storage locations on computing clusters and
clouds. CBRAIN offers transparent access to remote data sources,
distributed computing sites, and an array of processing and
visualization tools within a controlled, secure environment.  The
CBRAIN service deployed at the Montreal Neurological Institute relies
on the infrastructure provided by Compute Canada~\cite{das2016mni}. It
currently provides 500+ collaborators in 22 countries with web access
to several systems, including 6 clusters of the Compute Canada
high-performance computing infrastructure (totaling more than 100,000
computing cores and 40PB of disk storage) and Amazon EC2. CBRAIN
transiently stores about 10 million files representing over 50TB
distributed in 42 servers. 51 public data processing applications are
integrated and over 340,000 processing batches have been submitted since 2010.

Applications in CBRAIN are integrated as Ruby classes that create web
forms, validate parameters and run command lines on computing
resources. \boutiques is supported through a set of templates that
generate such classes from the application descriptor. Two application
integration modes are available:
\begin{enumerate}
  \item The descriptor is stored in a CBRAIN plugin and the Ruby classes
    are generated on-the-fly when CBRAIN starts. This mode allows
    CBRAIN developers update all \boutiques applications at once by
    editing the templates. However, it does not allow for
    customization beyond the \boutiques schema.
  \item Ruby classes are generated from descriptors
    through an offline process and integrated in CBRAIN as any other
    application. This mode allows developers to customize applications
    by editing the generated Ruby classes, but the resulting
    applications are difficult to maintain in the long term, in
    particular when the descriptors are updated.
\end{enumerate}

A custom property (\texttt{cbrain:inherits-from-class}) was added to
the \boutiques descriptor to define the Ruby class that should be used as
parent class for the application in CBRAIN. It provides more
flexibility to the first integration model described previously.

CBRAIN was also extended to support the parallelization mechanism
available in \boutiques: the CBRAIN execution agent periodically scans
the execution directories of the parallelized applications and it submits
the invocations defined in the JSON objects written by the
application in this directory.

A new list mechanism was also introduced in CBRAIN to facilitate the
iteration of \boutiques applications on large sets of files. CBRAIN
lists are specific files that contain references to other CBRAIN
files. When a list is passed to a \boutiques application, the elements
in the list are either concatenated in a single command line (when the
corresponding \boutiques input is a list), or a new command line is
generated for every element in the list (when the input is not a
list). Supporting lists as a specific CBRAIN file type allows for
improved validation. For instance, lists that contain references to
non-existent or deleted files can be detected. It also allows users to
edit lists using their own tools such as scripts or spreadsheet
applications.

\subsubsection{Flywheel}
%% gk: shall we check with them, and if nothing, remove?
\todo{Gunnar, Nathanael, John}

\subsubsection{Pegasus}
%% gk: to what extent does this exist?
\todo{Rafael}

\subsubsection{SPINE}
%% gk: shall we check with them, and if nothing, remove?
\todo{Charles}

\subsubsection{VIP}

The Virtual Imaging Platform (VIP)~\cite{GLAT-13} is a web portal for medical 
simulation  and image data analysis. VIP allows to make applications 
available as a service, and also to benefit from the resources available 
to the biomed Virtual Organisation (VO) of the European Grid Infrastructure 
\footnote{\url{https://www.egi.eu/}}, supported by approximately 65 computing
sites world-wide offering a total of 130 computing clusters and 5 PB of storage.
The VIP service is currently deployed at the Creatis laboratory
\footnote{\url{https://www.creatis.insa-lyon.fr/site7/en}} and uses the DIRAC 
French service \footnote{\url{https://dirac.france-grilles.fr}} to submit jobs on 
EGI grid and cloud resources.  VIP counts, in October 2017, more than 1000 
registered users and a growing number of available applications.
%% gk: I guessed 50 more in the last year, since 1000 is a nice number, is this accurate?

Until recently, applications were manually integrated in VIP as workflows written 
in the Gwendia\cite{MONT-09} language and managed  with the MOTEUR\cite{GLAT-08} workflow engine. 
As of today, Boutiques is supported through a "Boutiques application importer" tool, which 
parses the JSON file and automatically generates the corresponding application workflow
and the wrapper script that handles, among other things, the execution of the command line itself. 
The use of an application workflow allows to (i) iterate on input lists and generate 
parallel tasks, as well as to (ii) concatenate multiple tools. The applications used in 
the MICCAI challenges, for example, were followed by other tools allowing to compute 
challenge-specific metrics. Tool concatenation is currently handeled at the importer 
level based on pre-defined workflow templates.

\note{sorina}{@tristan: some info on workflows is available in the paragraph above. Is it OK/enough or do you have other suggestions?}

\subsection{Ported applications}

% cbrain-plugins-neuro

%%Table~\ref{table:applications} summarizes the applications described
%%with \boutiques and the main features used.

\paragraph{Niak.} The Niak pre-processing pipeline, executed with the Pipeline System
for Octave and Matlab (PSOM)~\cite{bellec2012pipeline}, was integrated
in CBRAIN through \boutiques. The integration uses the parallelization
mechanism described in Section~\ref{sec:parallelization} so that even
the invocations processing a single subject can be parallelized. It also
allows CBRAIN to leverage the efficient agent model used in PSOM, as
described in~\cite{GLAT-16}. The integration required some work in
PSOM to facilitate its invocation as a non-interactive command-line
application. The resulting CBRAIN plugin is available at
\url{https://github.com/SIMEXP/cbrain-plugins-psom}. It could in
principle be used in other platforms that support the parallelization
property, although this has not been tested.

\paragraph{Nipype.} The Nipype workflow engine~\cite{gorgolewski2011nipype} was
instrumented to export to \boutiques applications described in its
internal Python description language. \todo{Update nipype\_cmd to
  support the latest \boutiques schema and complete this paragraph.}

\paragraph{SPM.} The Statistical Parameter Mapping toolbox (SPM)~\cite{penny2011statistical}
was exported to Boutiques \ldots \todo{Guillaume}

\paragraph{MICCAI challenges.}
\boutiques was used to integrate 23 pipelines in the VIP platform in
the context of two challenges organized by the MICCAI conference in
2016, related to the segmentation of multiple-sclerosis lesions in MR
images (MSSEG
challenge\footnote{\url{https://portal.fli-iam.irisa.fr/msseg-challenge/overview}})
and of tumor volumes in PET images (PETSEG
challenge\footnote{\url{https://portal.fli-iam.irisa.fr/petseg-challenge/overview}}). The
pipelines were ported to VIP and executed on 205 subjects in a few
weeks only, which would not have been possible without
\boutiques. Some pipelines had to be adjusted manually once ported to
the platform for the following reasons:
\begin{enumerate}
\item A pipeline
  required a GPU, which we enabled through the
  \texttt{nvidia-docker}\footnote{\url{https://github.com/NVIDIA/nvidia-docker}}
  tool not supported in Boutiques by default although it could be a possible extension.
\item A pipeline required more than 10 GB of data dependencies (atlas
  data) which exceeded the maximum size allowed for Docker containers
  in our setup. We solved the issue by installing the data in a
  directory of the host server that we mounted in the container. Such
  platform-specific configurations could not possibly be supported by
  \boutiques.
\item A pipeline wrote more than 10 GB of intermediate data in a
  temporary directory of the container located on a 2 GB partition. We
  solved the issue by mounting a host directory in the temporary directory.
\end{enumerate}
The pipelines did not use much of the expressiveness features offered
by \boutiques since the instructions given to the challengers required
that all the pipeline parameters were hardcoded, to facilitate
execution. Two custom properties (\texttt{vip:miccai-challenger-email}
and \texttt{vip:miccai-challenge-team-name}) were added to the
\boutiques descriptor to help post-process results in the specific
context of MICCAI challenges.

\paragraph{ndmg.} The NeuroData MRI to Graphs one-click connectome estimation
pipeline~\cite{kiar2017comprehensive}, developed in Python and leveraging FSL, was
exported to \boutiques and is available at \url{https://github.com/neurodata/boutiques-tools}.
The ndmg pipeline is deployed in CBRAIN via its \boutiques descriptor, and is available
both through Docker and Singularity container environments.

\paragraph{BIDS apps.} Brain Imaging Data Structure supported
applications~\cite{gorgolewski2017bids}, a recent effort for the adoption of BIDS support
in common neuroimaging pipelines, require a standardized set of input and output parameters
by tools. A \boutiques tool has been developed which generates a descriptor for any such
BIDS app, enabling them to be deployed within a broad range of high performance computing
infrastructures, such as CBRAIN.

\paragraph{Other applications.} Other neuroinformatics applications were ported to CBRAIN using \boutiques:
FSL tools and ICA AROMA are available at
\url{https://github.com/aces/cbrain-plugins-neuro} and some pipelines
of the Human Connectome Project (HCP) are at
\url{https://github.com/big-data-lab-team/cbrain-plugins-hcp}.

\paragraph{VIP applications.} Several tools were imported in VIP using \boutiques. 
The first one is a phase contrast simulator called CreaPhase and developed at Creatis. 
The CreaPhase inputs had certain particularies (some nedeed to be enclosed 
in simple quotes, other were vectors of variable size enclosed in brackets)
that required the post-processing of the wrapper script generated by the
Boutiques importer.

The Odin MRI simulator\footnote{\url{http://od1n.sourceforge.net}} was also imported 
in VIP using \boutiques. Since Odin  may require important amounts of computing resources 
(the common use-case consists in iterating on a large number of inputs files), Odin is 
executed by VIP on the EGI grid resources, which currently do not support Docker. Consequently, 
the Odin executable was exported from the Docker image which allowed for his CentOS compilation
and the Odin wrapper script was modified so that it makes no "Docker run" call.

A few MRtrix3\footnote{\url{http://www.mrtrix.org/}} tools were also made available in VIP, as well as 
a script developed at Creatis, which combines MRtrix3 and FSL tools.

A few other applications from the FLI (France Life Imaging) consortium were also imported in
VIP using \boutiques. One of them is Ginfizz, a FMRI pipeline allowing to execute multiple 
preprocessing steps and basic analyses such as: (i) realignment (motion registration and correction),
(ii) slice timing (temporal interpolation to correct for the delay of acquisition between slices), 
(iii) coregistration (registering functional and structural data), (iv) normalization (registering 
of functional and anatomical data to a match a template that has to be specified), and (iv)
smoothing (spatial filtering to increase Signal to Noise ratio). The use of Docker containers
(\url{https://hub.docker.com/r/fliiam/ginfizz} based on \url{https://hub.docker.com/r/miykael/nipype_level2}) 
greatly simplifies its deployment. The animaN4BiasCorrection tool from the
Anima\footnote{\url{https://github.com/Inria-Visages/Anima-Public/wiki}} project was also made available 
in VIP. The import process was straightforward.

\note{sorina}{@tristan: please check if you agree with the descriptions above}

%%\begin{table}
%%  \begin{tabular}{l|ccccccc}
%%    \rowcolor[gray]{0.9}
%%    Application                      & \multicolumn{7}{c}{Input/Output properties} \\
%%    \rowcolor[gray]{0.9}
%%                                     & Dependencies      &Groups             & Lists             & Optional          &Default            & Type restriction             & Min/max\\
%%    
%%    \hline
%%    Niak pre-processing              &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\\
%%    ICA AROMA                        &\cellcolor{gray!75}&\cellcolor{gray!75}&                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
%%    FSL tools &&&&&&\\
%%    \multicolumn{1}{r|}{anat}        &\cellcolor{gray!75}&\cellcolor{gray!75}&                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
%%    \multicolumn{1}{r|}{bet}         &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&                   &                   &\cellcolor{gray!75}\\      
%%    \multicolumn{1}{r|}{fast}        &\cellcolor{gray!75}&                   &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
%%    \multicolumn{1}{r|}{first}       &                   &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&                   &                   \\
%%    FSL Nipype interfaces     \\
%%    HCP PreFreesurfer                &                   &                   &                   &                   &                   &                   &                   \\
%%    ndmg                &                   &       \cellcolor{gray!75}            &                   &          \cellcolor{gray!75}         &      \cellcolor{gray!75}             & \cellcolor{gray!75} &                   \\
%%    MICCAI challenge (23 pipelines)\\&                   &                   &                   &                   &                   &                   & \\
%%    \end{tabular}
%%  \vspace*{0.5cm}
%%  \begin{tabular}{l|ccc|ccc}
%%    \hline
%%    \rowcolor[gray]{0.9}
%%    Application                     &\multicolumn{3}{c|}{Containers}            &  \multicolumn{3}{c}{Other properties}  \\
%%    \rowcolor[gray]{0.9}
%%                                    &Docker              & Singularity& Rootfs  & Config. file & Env vars  & Parallelization   \\
%%    \hline
%%    Niak pre-processing             &\cellcolor{gray!75} &\cellcolor{gray!75}&         &              &\cellcolor{gray!75}&\cellcolor{gray!75}\\
%%    ICA AROMA                       &\cellcolor{gray!75} &            &         &              &&\\
%%    FSL tools &&&&&\\
%%    \multicolumn{1}{r|}{anat}       &                    &            &         &              &&\\
%%    \multicolumn{1}{r|}{bet}        &                    &            &         &              &&\\
%%    \multicolumn{1}{r|}{fast}       &                    &            &         &              &&\\
%%    \multicolumn{1}{r|}{first}      &                    &            &         &              &&\\
%%    FSL Nipype interfaces           &                    &            &         &              &&\\
%%    HCP PreFreesurfer               &\cellcolor{gray!75} &            &         &              &&\\
%%    ndmg                            &\cellcolor{gray!75} &\cellcolor{gray!75}&         &              &&\\
%%    MICCAI challenge (23 pipelines) &\cellcolor{gray!75} &            &         &              &&\\
%%  \end{tabular} 
%%  \caption{Applications available in Boutiques with descriptor features used.}
%%  \label{table:applications}
%%\end{table}

\section{Discussion}
With \boutiques, developers can port their applications once and execute
them in several platforms. \boutiques removes the technological dependency to a
particular platform and facilitates migration.  Although the
motivating use cases were taken from neuroinformatics, our primary
field of interest, nothing prevents the system from being used in a
variety of other domains.

\subsection{\boutiques descriptor}

The \boutiques descriptor specification allows describing a wide range
of applications, but it is also getting increasingly complex through
additions such as invocation schemas and dependencies among
inputs. Extending the \boutiques descriptor has two main goals. First,
accurately describing applications improves validation: incorrect input
values and execution results are more precisely detected when the
application descriptor is comprehensive, which is essential to improve
the reliability and user-friendliness of execution
platforms. Secondly, a rich descriptor schema reduces the need for
custom wrappers to integrate applications as even convoluted command
lines can be described: this is particularly interesting for
applications installed in containers since adding a wrapper to the
execution stack requires an update in the container image.

Nonetheless, a complex descriptor schema has a cost for application
developers and platforms, which we address as follows. For developers,
we maintain the set of mandatory descriptor properties as small as
possible so that simple applications can be described in a few lines
only (see Listing~\ref{listing:minimal}). For platforms, we aim at
supporting as many features as possible in the executor so that
only the following steps need to be implemented to support \boutiques
in a platform, regardless of the complexity of the descriptor:
\begin{enumerate}
  \item Input entry: generate the interface to allow users to enter
    input values.
  \item Input validation: create a JSON object from the interface,
    validate it against the invocation schema.
  \item Execution: generate and run the command line from the local
    executor and input JSON object.
  \item Output delivery: from the descriptor, identify the output files
    and deliver them to the user.
\end{enumerate}
In particular, advanced validation features such as dependencies
between inputs are embedded in the regular JSON validation performed
at step 2, without requiring the platform to support the related
descriptor properties.
\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=21pt,
               tabsize=4]{js}
{
    "name": "echo",
    "tool-version": "1.0",
    "description": "A simple script to test output files",
    "command-line": "echo [PARAM] > output.txt",
    "schema-version": "0.4",
    "inputs": [{
	"id": "param",
	"name": "Parameter",
        "value-key": "[INPUT_FILE]",
	"type": "File"
    }],
    "output-files": [{
	"id": "output_file",
	"name": "Output file",
	"path-template": "output.txt"
    }]

}
\end{minted}
\caption{A minimal \boutiques descriptor.} 
\label{listing:minimal}
\end{listing}

\subsection{Workflow support}

The parallelization framework described in~\ref{sec:parallelization}
supports workflows by considering them as both applications and
platforms, which we think is a powerful but underexploited
model. Workflows are applications because they are invoked like any
other script or executable, possibly through a command line. Workflows
are also platforms because they may invoke other applications that
might themselves be described in \boutiques. This model is powerful
because it shields the \boutiques specification from specific workflow
constructs and allows a wide range of workflow engines to be described
uniformly.

\subsection{Reproducibility}

\boutiques helps reproduce\footnote{in Peng's
  sense~\cite{peng2011reproducible}} executions through containers and
formal command-line descriptions. With \boutiques, complete sets of
applications could be easily migrated across execution platforms,
including high-performance computing platforms and individual laptops,
to reproduce analyses. The \boutiques descriptor describes the
parameters and implementation of the application, and the invocation
schema describes the parameter values.

However, reproducibility is a large problem that \boutiques only
partially addresses. At the command-line execution level, containers
help freeze a large fraction of the software ecosystem but they do not
shield against discrepancies arising from different Linux kernel
versions or hardware platforms. For instance, containers may not
execute consistently on different x86\_64 CPUs (e.g. Intel and AMD)
when software compiled with architecture-specific flags is involved
(see for instance GCC's \texttt{-march} flag).

In addition, important runtime parameters, for instance related to
multi-threading or available resources (storage, RAM), may be set by
the execution platform without being specified in the \boutiques
descriptor. Such runtime parameters may impact reproducibility in some
cases. To properly cover this issue, \boutiques descriptors should be
complemented by a provenance framework that captures a detailed trace
of the execution.

At a higher level, reproducibility is impacted by factors
such as data sharing~\cite{poldrack2014making} which are out of the
scope of \boutiques.

\subsection{Application types}

So far, \boutiques has focused on the description of non-interactive
command-line applications. While such applications cover a large
fractions of the tools involved in scientific data processing, other
types of programs exist such as web services, interactive tools, and
applications with a graphical user interface (GUI).  Such application
classes could be described in \boutiques through a command-line
mapping. For instance, web services may be wrapped as command-line
applications using tools such as \texttt{curl} or
\texttt{wget}. Interactive applications may also be transformed to
non-interactive ones through configuration files. Finally, nothing
prevents a \boutiques tool from popping up a window for a user to
provide input through a GUI. This should, however, be specified
as an extension to the descriptor since most platforms would not support
this feature by default. Graphical output produced by applications
executed in containers should also be treated specifically.

\subsection{Limitations}

A few limitations remain that should be addressed in the
future. First, \boutiques moves the application porting bottleneck
from integration to validation. Using \boutiques, functions can be
automatically exported from frameworks such as Nipype and SPM,
creating hundreds of richly described applications potentially usable
by end-users. However, it remains unclear how to validate that these
applications actually execute as expected beyond executing packaged tests
distributed alongside a given tool. A \boutiques-specific testing
framework could be designed and potentially fed by existing frameworks,
to address this issue.

Another limitation is related to the security of containerized
applications. Since containers are usually controlled by application
developers rather than platforms, which is a good thing to reduce
application porting bottlenecks, nothing prevents developers to embed
malicious code in their container at any stage of the process,
possibly after a platform administrator inspected the
container. Containers are currently bulky file archives that are
cumbersome to inspect. Tools need to be developed to allow for an
easier characterization of container contents, for instance through
comparison digests with respect to validated base images. Singularity
containers have reduced security risks as compared to Docker containers,
but the issue of content transparency is still not avoided.

% How to build modular container images for workflows.

\section{Related work}

The Common Workflow Language\footnote{\url{http://www.commonwl.org}}
is the piece of related work most closely related to
\boutiques. Several other application description frameworks have been
developed but they do not support containers.

\subsection{Common Workflow Language}

The Command Line Tool Description of the Common Workflow Language
(CWL) overlaps with the \boutiques descriptor. According to the history
of the CWL and \boutiques Github repositories, CWL started 6 months
before \boutiques (September 2014 vs. May 2015). The differences
highlighted below were identified from version 1.0 of the CWL Command
Line Tool
Description\footnote{\url{http://www.commonwl.org/v1.0/CommandLineTool.html}}.

\subsubsection{Conceptual differences}

The following differences are conceptual in the sense that they may
not be easily addressed by CWL or \boutiques developers without deeply
refactoring the frameworks.

First, CWL has a workflow language whereas \boutiques does not. In
\boutiques, workflows are integrated as any other applications, except
that they may submit other invocations to enable workflow
parallelism. This fundamental difference has consequences on the
complexity of CWL application descriptions and on the possibility to
reuse existing workflows in \boutiques. The adoption of ontologies in
CWL may also be another consequence (see below).

CWL has a formal command-line definition while \boutiques is more
flexible. CWL specifies command lines using an array containing an
executable and a set of arguments whereas \boutiques only uses a
string template. \boutiques' template approach may create issues in
some cases, but it also allows developers to add simple operations to
an application without having to write a specific wrapper. For
instance, a \boutiques command line may easily include input
decompression using, e.g., the tar command in addition to the main
tool command. On the contrary, CWL applications may only have a single
executable. Importantly, \boutiques' template system allows supporting
configuration files (see Section~\ref{sec:configuration-files}).

CWL uses ontologies while \boutiques does not. Ontologies allow for
richer definitions but they also have an overhead. The main
consequences are the following:
\begin{itemize}
\item CWL uses a specific framework for
validation, called SALAD (Semantic Annotations for Linked Avro Data)
whereas \boutiques uses plain JSON schema. The main goal of SALAD is to
allow ``working with complex data structures and document formats, such
as schemas, object references, and namespaces''. \boutiques only relies
on the basic types required to describe and validate a command line
syntactically. While the use of SALAD certainly allows for
higher-level validation, and may simplify the composition and
validation of complex workflows, it also introduces a substantial
overhead in the specification, and platforms have to use the validator
provided by CWL. On the contrary, a regular JSON validator can be used in \boutiques.
\item  CWL has a rich set of types whereas \boutiques only has simple
types. Again, this may be seen as a feature or as an overhead
depending on the context. \boutiques tries to limit the complexity of
the specification to facilitate its support by platforms where tools
will be integrated.
\end{itemize}

\subsubsection{Major differences}

The following differences are major but they may be addressed by the
CWL and \boutiques developers as they do not undermine the application
description model.

\begin{itemize}
\item CWL tools have to write in a specific set of directories called
``designated output directory'', ``designated temporary directory''
and ``system temporary directory''. Tools are informed of the location
of such directories through environment variables. Having to write in
specific directories is problematic because tools have to be modified
to enable that. In \boutiques, the path of output files is defined
using a dedicated property.
\item CWL types are richer, not only
semantically but also syntactically. For instance, files have
properties for basename, dirname, location, path, checksum, etc.
\item \boutiques supports various types of containers (Docker, Singularity,
rootfs) while CWL supports only Docker. Both tools have rich requirements:
for instance, they may include RAM, disk usage, and walltime estimate; 
whereas \boutiques also supports number of requested CPU cores. CWL has hints, i.e.,
recommendations that only lead to warnings when not respected.
\item In \boutiques, dependencies can be defined among inputs, e.g., to
specify that an input may be used only when a particular flag is
activated. This is a very useful feature to improve validation, in
particular when tools have a lot of options.
\item In \boutiques, named
groups of inputs can be defined, which improves the presentation of
long parameter lists for the user and enables the definition of more
constraints within groups (e.g. mutually exclusive inputs).
\end{itemize}

\subsubsection{Other minor differences}

The \boutiques and CWL schemas have other minor differences that we
summarized in a public
spreadsheet\footnote{\url{https://docs.google.com/spreadsheets/d/1eLhpxzaaMPCIUFmV5Trs-WtaMpBO-c6QDQsmhwP7QIE/edit\#gid=0}}.

\subsection{Other frameworks}

Several frameworks have been created to facilitate the
integration of command-line applications in platforms. In
neuroinformatics, many platforms define a formal interface to embed
command-line applications. Among them, the Common
Toolkit\footnote{\url{http://www.commontk.org}} interoperates with
several platforms such as 3D Slicer~\cite{pieper20043d},
NiftyView~\cite{Craddock2016}, GIMIAS~\cite{larrabide2009gimias},
MedInria~\cite{larrabide2009gimias}, MeVisLab~\cite{heckel2009object}
and MITK workbench~\cite{nolden2013medical}. The framework, however,
remains tightly bound to the Common Toolkit's C++ implementation which
limits its adoption, e.g., in web platforms \todo{Fr\'ed\'eric, maybe
  you have more to say on this}.

In the distributed computing community, systems were also proposed to
facilitate the embedding of applications in platforms. The Grid
Execution Management for Legacy Code Architecture
(GEMLCA~\cite{delaitre2005gemlca}) was used to wrap applications in
grid computing systems. Interestingly, it has been used to embed
workflow engines in the SHIWA
platform~\cite{terstyanszky2014enabling}, in a similar but different
way than proposed by \boutiques (see Section~\ref{sec:parallelization}).

The recent advent of virtual containers requires a new generation of
application description frameworks that are independent from any
programming language and that expose a rich set of properties to
describe command lines, as intended by \boutiques.

\section{Conclusion}

\boutiques is available at \url{https://github.com/boutiques}. We
welcome feedback, issue reporting and pull requests. \boutiques adopts
a bottom-up approach where new features are progressively added based
on feedback from applications and platforms. Beyond the technicalities
discussed in this paper, the availability of a solid core of
applications and platforms in the framework is key to its success,
which we plan to continuously enhance.

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}