### 2018-02-01 letter.tex

\documentclass[a4]{letter}
\usepackage{graphicx}
\name{Tristan Glatard}
\signature{\includegraphics[width=4cm]{signature.png}}
\date{January 26th, 2018}

\begin{document}

\begin{letter}{}

\opening{Dear Editors,}

We would like to thank you for offering us the opportunity to revise
and re-submit our paper entitled ``Boutiques: a flexible framework to integrate
command-line applications in computing platforms'' to GigaScience.

The remainder of this letter contains our answer to the 2 reviews that
we received following the evaluation process. A revised version of the
manuscript follows.

We thank you again for your time and consideration, and we remain
available for any further information you may need.

\vspace{0.5cm}

Sincerely yours,

\vspace{1cm}

\closing{Tristan Glatard, PhD.\\On behalf of the authors.}

\end{letter}
\end{document}